Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-07-01 optim.qmd
revised section 4.2

In addition to the pipe-finding problem, a new shape, sine wave, is investigated in 6D and 8D spaces with six indexes considered:
`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. Combining with two JSO hyper-parameters, a total of `r nrow(sim_df)` cases is produced, comprising of `r nrow(filter(sim_df, index == "holes"))` pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))` sine-wave finding cases. For each case, JSO is run fifty times and the success rate is calculated.

Smoothness and squintability are computed for each case, following the procedures outlined in @sec-smoothness and @sec-squintability and illustrated in @fig-smoothness and  @fig-squintability. For smoothness, three hundred random bases are simulated. Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting into a Gaussian process model to obtain 
the smoothness measure. For squintability, fifty random bases are sampled and interpolated to the optimal basis with a step size of 0.005. Index values and projection 
distances are calculated for these interpolated bases and the index values are averaged with a bin width of 0.005. The squintability measure is then calculated from fitting a 4-parameter scaled logistic
function of index value against projection distance, estimated by non-linear least squares.

Combining with the success rate calculated for each case, the dataset is fitted using  a generalised linear model with a binomial family and a logit link
function with three pre-processing steps are applied: 
1) scaling the JSO hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) encoding the success rate for
the stringy case as 0, because upon visual inspection of the final projection, none of the 50 simulations identified the sine-wave shape.

@@

```{r}
#| label: tbl-mod-data
#| tbl-cap: The first 7 rows of the datasets processed for modelling. 
sim_df |> 
  dplyr::select(index, d, smoothness, 
    squintability, n_jellies, max_tries, 
    -I_max, P_J, time) |> 
  mutate(time = as.numeric(time)) |>
  arrange(index, d, n_jellies, max_tries) |> 
  head(7) |> 
  kable(digits = 2, format = "latex", 
        col.names = c("Index", "D", 
          "Smoothness", "Squintability", 
          "N. jellyfish", 
          "Max. tries", 
          "Success rate", "Time (sec)"),
        linesep = "",
        booktabs = T) |> 
  kable_styling(#latex_options = "hold_position",
    font_size = 10)
```

>>>>

This simulation uses two different data shapes, pipe and a sine wave, investigated in 6D and 8D spaces with six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. The two JSO hyper-parameters are varied. A total of `r nrow(sim_df)` combinations result, comprising of `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and `r nrow(filter(sim_df, index != "holes"))` on the sine-wave data. For each case, JSO is run 50 times to calculate the success rate.

Smoothness and squintability are computed for each run, following the procedures outlined in @sec-smoothness and @sec-squintability and as illustrated in @fig-smoothness and  @fig-squintability.

To compute smoothness, 300 random bases are simulated. Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting the Gaussian process model to obtain the smoothness measure for the index.

To compute squintability, 50 random bases are sampled and interpolated to the optimal basis with a step size of 0.005. Index values and projection distances are calculated for these interpolated bases and the index values are averaged with a bin width of 0.005. The squintability measure is then calculated by fitting a four-parameter scaled logistic function of index value against projection distance, estimated by non-linear least squares.

A generalised linear model is fitted to the resulting data, using a binomial family and a logit link function. The data is pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) re-coding the success rate for
the `stringy` index as 0, because none of the 50 simulations correctly identified the sine-wave shape.

@@

```{r}
#| label: tbl-mod-data
#| tbl-cap: The first 7 rows of the datasets processed for modelling. 
sim_df |> 
  dplyr::select(index, d, smoothness, 
    squintability, n_jellies, max_tries, 
    -I_max, P_J, time) |> 
  mutate(time = as.numeric(time)) |>
  arrange(index, d, n_jellies, max_tries) |> 
  head(7) |> 
  kable(digits = 2, format = "latex", 
        col.names = c("index", "d", 
          "smoothness", "squintability", 
          "n. jellyfish", 
          "max. tries", 
          "success rate", "time (sec)"),
        linesep = "",
        booktabs = T) |> 
  kable_styling(#latex_options = "hold_position",
    font_size = 10)
```

### 2024-07-01 optim.tex
revised section 4.2

\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, opacityback=0, breakable, toprule=.15mm, arc=.35mm, colback=white, coltitle=black, bottomrule=.15mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, rightrule=.15mm, opacitybacktitle=0.6]

@@

In addition to the pipe-finding problem, a new shape, sine wave, is
investigated in 6D and 8D spaces with six indexes considered:
\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}. Combining with two JSO
hyper-parameters, a total of 52 cases is produced, comprising of 24
pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
run fifty times and the success rate is calculated.

Smoothness and squintability are computed for each case, following the
procedures outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability} and illustrated in
Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. For
smoothness, three hundred random bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting into a Gaussian process model to obtain the
smoothness measure. For squintability, fifty random bases are sampled
and interpolated to the optimal basis with a step size of 0.005. Index
values and projection distances are calculated for these interpolated
bases and the index values are averaged with a bin width of 0.005. The
squintability measure is then calculated from fitting a 4-parameter
scaled logistic function of index value against projection distance,
estimated by non-linear least squares.

Combining with the success rate calculated for each case, the dataset is
fitted using a generalised linear model with a binomial family and a
logit link function with three pre-processing steps are applied: 1)
scaling the JSO hyper-parameters by a factor of 10 for interpretation,
2) creating a new binary variable \texttt{long\_time} to indicate cases
with an average run time over 30 seconds, and 3) encoding the success
rate for the stringy case as 0, because upon visual inspection of the
final projection, none of the 50 simulations identified the sine-wave
shape.

@@

\begin{tabular}{lrrrrrrr}
\toprule
Index & D & Smoothness & Squintability & N. jellyfish & Max. tries & Success rate & Time (sec)\\
\midrule
MIC & 6 & 2.46 & 1.26 & 20 & 50 & 0.12 & 2.48\\
MIC & 6 & 2.46 & 1.26 & 20 & 100 & 0.24 & 8.95\\
MIC & 6 & 2.46 & 1.26 & 50 & 50 & 0.52 & 5.65\\
MIC & 6 & 2.46 & 1.26 & 50 & 100 & 0.64 & 13.22\\
MIC & 6 & 2.46 & 1.26 & 100 & 50 & 0.76 & 19.45\\
MIC & 8 & 2.64 & 0.77 & 20 & 50 & 0.08 & 2.57\\
MIC & 8 & 2.64 & 0.77 & 20 & 100 & 0.08 & 4.96\\
\bottomrule
\end{tabular}
\endgroup{}

>>>>

\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, coltitle=black, colback=white, colbacktitle=quarto-callout-note-color!10!white, left=2mm, leftrule=.75mm, titlerule=0mm, rightrule=.15mm, toptitle=1mm, opacityback=0, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, opacitybacktitle=0.6, breakable]

@@

This simulation uses two different data shapes, pipe and a sine wave,
investigated in 6D and 8D spaces with six different PP indexes:
\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}. The two JSO hyper-parameters are
varied. A total of 52 combinations result, comprising of 24 computed on
the pipe data and 28 on the sine-wave data. For each case, JSO is run 50
times to calculate the success rate.

Smoothness and squintability are computed for each run, following the
procedures outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability} and as illustrated in
Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}.

To compute smoothness, 300 random bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting the Gaussian process model to obtain the
smoothness measure for the index.

To compute squintability, 50 random bases are sampled and interpolated
to the optimal basis with a step size of 0.005. Index values and
projection distances are calculated for these interpolated bases and the
index values are averaged with a bin width of 0.005. The squintability
measure is then calculated by fitting a four-parameter scaled logistic
function of index value against projection distance, estimated by
non-linear least squares.

A generalised linear model is fitted to the resulting data, using a
binomial family and a logit link function. The data is pre-processed by
1) scaling the JSO hyper-parameters by a factor of 10 for
interpretation, 2) creating a new binary variable \texttt{long\_time} to
indicate cases with an average run time over 30 seconds, and 3)
re-coding the success rate for the \texttt{stringy} index as 0, because
none of the 50 simulations correctly identified the sine-wave shape.

@@

\begin{tabular}{lrrrrrrr}
\toprule
index & d & smoothness & squintability & n. jellyfish & max. tries & success rate & time (sec)\\
\midrule
MIC & 6 & 2.46 & 1.26 & 20 & 50 & 0.12 & 2.48\\
MIC & 6 & 2.46 & 1.26 & 20 & 100 & 0.24 & 8.95\\
MIC & 6 & 2.46 & 1.26 & 50 & 50 & 0.52 & 5.65\\
MIC & 6 & 2.46 & 1.26 & 50 & 100 & 0.64 & 13.22\\
MIC & 6 & 2.46 & 1.26 & 100 & 50 & 0.76 & 19.45\\
MIC & 8 & 2.64 & 0.77 & 20 & 50 & 0.08 & 2.57\\
MIC & 8 & 2.64 & 0.77 & 20 & 100 & 0.08 & 4.96\\
\bottomrule
\end{tabular}
\endgroup{}
